Peer Assessment 1 for course "Reproducible Research"
========================================================

Load data from CSV file:
```{r load.data}
data <- read.csv("activity.csv", header=TRUE, stringsAsFactors=FALSE)
```

Convert variable "date" to proper R date format:
```{r convert.date}
data$date <- as.Date(strptime(data$date, format="%Y-%m-%d"))
```

Analysis of mean total number of steps taken per day, for this we omit
incomplete measurements:
```{r drop.na}
comp.data <- data[complete.cases(data),]
```

Calculate number of steps per day and assign meaning variables names:
```{r calc.daily.steps}
daily.steps <- aggregate(comp.data$steps,list(comp.data$date),FUN=sum)
col.names <- c("date","total.steps")
colnames(daily.steps) <- col.names
```

Plot total number of steps per day:
```{r fig.width=10, fig.height=6}
hist(daily.steps$total.steps,breaks=10,col="red",main="total number of steps per day",xlab="number of steps per day",ylab="frequency")
```

Calculate mean and median of total number of steps per day:
```{r calc.mean.median}
paste("Mean value of total number of steps per day:",format(mean(daily.steps$total.steps),digits=5))
paste("Median value of total number of steps per day:",median(daily.steps$total.steps))
```

Calculate average number of steps taken per 5-minute interval:
```{r interval.averages}
interval.avg <- aggregate(comp.data$steps,list(comp.data$interval),FUN=mean,na.omit=TRUE)
col.names <- c("interval.no","avg.steps")
colnames(interval.avg) <- col.names
```

Plot average no of steps per interval:
```{r fig.width=10, fig.height=6}
plot(interval.avg,type="l",main="average no of steps per 5-minute interval",xlab="interval number","average no of steps")
```

Find interval with maximum number of steps:
```{r max.interval}
max.int <- sort(interval.avg$avg.steps,decreasing=TRUE,index.return=TRUE)
paste("Index of interval with maximum steps (daily average):",max.int$ix[1])
```

Calculate total number of rows with missing values:
```{r missing.values}
paste("Number of rows with missing values:",nrow(data)-nrow(comp.data))
```
